SMT-Boosted Security Types for Low-Level MPC

### 2025-11-23 toplas26/sty.tex
checkpoint

The \TirName{Share} rule generates the binding $[\mx{z}{\cid_2},
  \rx{y}{\cid_1}] : \sharety$ if this is a secret sharing as defined
by $\genshares$. It also adds $\rx{y}{\cid_1}$ to the set of variables
consumed by $\genshares$, ensuring that it has not already been used. The
\TirName{Reveal} rule generates the binding $\pubx{\recon(\px{x},\px{y})} :
\outty$ with the requirement that $[\px{x},\px{y}]$ are typable as
shares. The \TirName{SecOpen} rule is similar except it allows the
assertion $\pubx{\recon(\px{x},\px{y})} : \pubty$ if we can additionally
show that the shared value is masked by the use of random value
$\rx{z}{\oid}$ as an OTP. Here we are positing a third party $\oid$
that secret-shares $\rx{z}{\oid}$ as part of an initialization
protocol. We illustrate secure openings in practice at more length
in Section \ref{section-examples}, though we note that appropriate
properties of initialization protocols can be encoded in $\eqspre$.

@@

To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
intuitively meaning that all the share types in $\Gamma$ are sound with respect
to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
Property (i) ensures that views do not reveal information about secrets.
Property (ii) ensures the key property of sharing-- that individual shares
don't reveal more about secrets than their reconstruction does. Properties
(iii) and (iv) ensure that OTPs used for generating shares or for
secure openings, respectively, are included in $R$.
\begin{definition}
  \label{def-OK}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    $.
  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}
With the definition of ${OK}$ in place, we can define what we mean by soundness
of $\genshares$. In this definition, the first condition establishes that
initial shares are independent of honest secrets, and the second establishes
that they reveal no more than their reconstruction. These conditions are
obviously met by ``vanilla'' additive or Shamir shares which generate
shares in uniform marginal distributions.
\begin{definition}
  \label{def-genshares}
  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
  with $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  we have for all $a,b \in \setit{1,2}$:
  \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
    \end{enumerate}
\end{definition}

Next we show that, given well-formed typing environments, share type judgements preserve
the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
known semi-homomorphic properties of non-interactive operations that are syntactically
verified in typing judgements, and by Definition \ref{def-genshares} and the
assumption of soundness of $\genshares$.
\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
       \setit{\phi_1,\ldots,\phi_j} \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1, \phi_i^2), \phi_i^b}}$
    \end{enumerate}
\end{lemma}
Next, we show that interactive typing rules preserve well-formedness of type
environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound}
and the security of share generation using OTPs-- that is, shares
generated by $\genshares$ are in uniform marginal distributions. 
\begin{lemma}
  \label{lemma-OK}
  Given $\prog,\Gamma_1,R_1  : {OK}$
  and $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
  Then $\toeq{\prog}, \Gamma_2, R_2 : {OK}$.
\end{lemma}
In addition to typing judgements, in order to show that a protocol is
secure, we need to classify all corrupt views as shares, all
public values as secure openings, and all outputs as reconstructed
shares. We capture this formally with the following notion of
type environment closure.
\begin{definition}
  \label{def-sty-closed}
  Given  $\prog,\Gamma,R  : {OK}$ and $\iov(\prog) = S \cup V \cup O$.
  Then $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
  for all $x \in \houtputs$ there exists $\phi$ such that either
  $\Gamma,\toeq{\prog} \vdash [x,\phi] : \sharety$ if $C = \setit{1}$
  or  $\Gamma,\toeq{\prog} \vdash [\phi,x] : \sharety$ if $C = \setit{2}$,
  and for all $\out{\cid} \in O$ we have $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$.
\end{definition}
Now, the following is a consequence of Definitions \ref{def-OK} and \ref{def-sty-closed}, and then
our main Theorem \ref{theorem-sty} follows by Lemma \ref{lemma-OK} and Theorem \ref{theorem-OK}.
\begin{theorem}
  \label{theorem-OK}
  If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}
\begin{theorem}
  \label{theorem-sty}
  If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}

>>>>

The \TirName{GenShares} rule generates the binding $[x_1,
  x_2] : \sharety$ if the term is a secret sharing as defined
by $\genshares$, and adds $\rx{y}{\cid_1}$ to the set of variables
consumed by $\genshares$, ensuring one-time usage. The
\TirName{Reveal} rule generates the binding $\pubx{\recon(\px{x},\px{y})} :
\outty$ with the requirement that $[\px{x},\px{y}]$ are typable as
shares. The \TirName{SecOpen} rule is similar except it allows the
assertion $\pubx{\recon(\px{x},\px{y})} : \pubty$ if we can additionally
show that the shared value is masked by the use of random value
$\rx{z}{\oid}$ as an OTP. Here we are positing a third party $\oid$
that secret-shares $\rx{z}{\oid}$ as part of an initialization
protocol. We illustrate secure openings in practice at more length
in Section \ref{section-examples}, though we note that appropriate
properties of initialization protocols can be encoded in $\eqspre$.

@@

To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
intuitively meaning that all the share types in $\Gamma$ are sound with respect
to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
Property (i) ensures that views do not reveal information about secrets.
Property (ii) ensures the key property of sharing-- that individual shares
don't reveal more about secrets than their reconstruction does. Property
(iii) ensures that public values are independent of outputs or secrets. Properties
(iv) and (v) ensure that OTPs used for generating shares or for
secure openings, respectively, are included in $R$.
\begin{definition}
  \label{def-OK}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    $.
  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
  \item $\sep{\progd(\prog)}{(S \cup O)}{\setit{\phi_1,\ldots,\phi_j}}$
  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}
With the definition of ${OK}$ in place, we can define what we mean by soundness
of $\genshares$. In this definition, the first condition establishes that
initial shares are independent of honest secrets, and the second establishes
that they reveal no more than their reconstruction. These conditions are
obviously met by ``vanilla'' additive or Shamir shares which generate
shares in uniform marginal distributions.
\begin{definition}
  \label{def-genshares}
  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
  with $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  we have for all $a,b \in \setit{1,2}$:
  \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
    \end{enumerate}
\end{definition}

Next we show that, given well-formed typing environments, share type judgements preserve
the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
known semi-homomorphic properties of non-interactive operations that are syntactically
verified in typing judgements.
\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
      % \setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1, \phi_i^2), \phi_i^b}}$
    \end{enumerate}
\end{lemma}
\begin{proof}
  The proof is by induction on the derivation of $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$.
  In the base case where the judgement follows by the \TirName{Shared} rule, the result
  follows trivially by Definition \ref{def-OK}. In cases \TirName{HEAdd}, \TirName{HEAddPub}, and
  \TirName{HEMultPub} the result follows by Definition \ref{def-OK} and known
  semi-homorphic properties of secret sharing \cite{10.1007/978-3-030-68869-1_3}.
  Case \TirName{ShareEntails} follows by correctness of the SMT interpretation of
  protocols as in Theorem \ref{theorem-toeq}.
\end{proof}

Next, we show that interactive typing rules preserve well-formedness of type
environments. This Lemma follows by Lemma \ref{lemma-sty-noninteractive-sound},
and the security of share generation using OTPs as articulated in
Definition \ref{def-genshares}-- that is, shares
generated by $\genshares$ are in uniform marginal distributions. 
\begin{lemma}
  \label{lemma-OK}
  Given $\prog,\Gamma_1,R_1  : {OK}$
  and $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
  Then $\toeq{\prog}, \Gamma_2, R_2 : {OK}$.
\end{lemma}
\begin{proof}
  The proof follows by case analysis on  $\toeq{\prog}  \vdash \Gamma_1, R_1 \stredx \Gamma_2, R_2$.
  In case \TirName{GenShares} the result follows by Definition \ref{def-genshares},
  which satisfies relevant conditions (i-ii,iv) of Definition \ref{def-OK}. In case
  \TirName{SecOpen} the public value must be reconstructed shares so Lemma
  \ref{lemma-sty-noninteractive-sound} relevant conditions (i-ii), and the
  the requirement that the public value be masked with an OTP satisfies conditions
  (iii) and (iv) of Definition \ref{def-OK}. In case \TirName{Reveal} since
  the output value must be reconstructed shares, Lemma \ref{lemma-sty-noninteractive-sound}
  satisfies relevant case (ii) of  Definition \ref{def-OK}. 
\end{proof}

In addition to typing judgements, in order to show that a protocol is
secure, we need to classify all corrupt views as shares, all
public values as secure openings, and all outputs as reconstructed
shares. We capture this formally with the following notion of
type environment closure.
\begin{definition}
  \label{def-sty-closed}
  $\Gamma$ is \emph{closed for $\prog$} iff for all $H,C$,
  for all $x \in \houtputs$ there exists $\phi$ such that either
  $([x,\phi] : \sharety) \in \Gamma$ if $C = \setit{1}$
  or  $([\phi,x] : \sharety) \in \Gamma$ if $C = \setit{2}$,
  and for all $\out{\cid} \in O$ we have $(\pubx{\out{\cid}} : \outty) \in \Gamma$.
\end{definition}
Now, the following is a consequence of Definitions \ref{def-OK} and \ref{def-sty-closed}, and then
our main Theorem \ref{theorem-sty} follows by Lemma \ref{lemma-OK} and Theorem \ref{theorem-OK}.
\begin{theorem}
  \label{theorem-OK}
  If $\prog,\Gamma,R : {OK}$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}
\begin{proof}
  Let $\iov(\prog) = S \cup V \cup O$ and assume some partitioning $H,C$.
  %For any $\out{\cid} \in O$ with $\Gamma,\toeq{\prog} \vdash \pubx{\out{\cid}} : \outty$
  %we can assume wlog that $(\out{\cid} : \outty) \in \Gamma$, since we can
  %add the binding to $\Gamma'$ with  $\prog,\Gamma',R  : {OK}$ by Lemma
  %\ref{lemma-OK}.
  By Definitions \ref{def-sty-closed} we know that every $x \in \houtputs$
  occurs in some $\stt$ with $(\stt, \sharety) \in \Gamma$, and
  that $(\out{\cid} : \outty) \in \Gamma$ for all $\out{\cid} \in O$.
  Therefore by Definition \ref{def-OK} (and marginalizing away
  extraneous $(\stt : \sty) \in \Gamma$) we have:
  $$
  \condd{\progd(\prog)}{S_H}{S_C \cup O} = \condd{\progd(\prog)}{S_H}{S_C \cup V_C \cup O}
  $$
  which is noninterference modulo output by Definition \ref{definition-NIMO}.
\end{proof}

\begin{theorem}
  \label{theorem-sty}
  If $\toeq{\prog} \vdash \varnothing,\varnothing \stredx \Gamma,R$ and $\Gamma$ is closed
  for $\prog$, then $\prog$ satisfies noninterference modulo output. 
\end{theorem}

\begin{proof}
  By Lemma \ref{lemma-OK} we have that $\prog,\Gamma,R : {OK}$, so the result follows
  by Therorem \ref{theorem-OK}.
\end{proof}
